---
title: "FinalProject"
author: "Arash Zakeresfahani"
date: "12/18/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import these packages:
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(readr)
library(ngram)
library(dplyr)
library(stringr)
library(prob)

```

Import/read text file

```{r}
dt = read_file('98-0.txt')
```

This function takes the text from the first chapter of "A TALE OF TWO CITIES" book from homework one. I had to limit it to the first chapter because of memory limits, otherwise any cleaned text works with the following functions(need to comment out a line in the "predic" function in order to work with other texts).
It returns a the first chapter of the book without control charecters and parentesis 
```{r}
#text_book = read_file('98-0.txt')

clean_txt = function(txt){
  dt = unlist(regmatches(txt,gregexpr("I\\. The Period.*(II\\. The Mail)", txt)))
  dt <- dt %>% str_replace_all( "[[:cntrl:]]", " ") %>%
     str_replace_all("\\s+", " ")%>%
     str_replace_all("II\\. The Mail", " ") %>%
     str_replace_all("(\\(|\\))+", " ")
  return(dt)
}
```

This function create a probablity matrix. First, we create a matrix of 0 with all of the words added as row and column names and add ones to each next word diagnaly. 

Second, I get the sum of each row between the duplicated column names and assign the values to a new matrix 
(uni_col_mat) with all the words as row names and only unique words as column names. 

Third, I get the sum between values for each column between the rows with duplicated names and add the values to new matrix (uni_row_mat). The uni_row_mat has unique words as row and column names. The values in the uni_row_mat represet the  all frequencies a word(column name) coming after the previous word(row name). 

Next, we convert the frequencies to probability fractions. It returns a probablity matrix. 
```{r}
prob_mat = function(txt){
  
  # Split the words by space
  words = unlist(strsplit(txt, " "))
  
  #Drop the empty charecter
  words = words[words != ""]
  
  #Create matrix of zeros with dimansions = number of words in the text and all of the words as row and column names. 
  all_mat = matrix(ncol = length(words), nrow = length(words),0)
  #Use the words for rows and columns names
  colnames(all_mat) = words
  rownames(all_mat) = words

  # Add 1 to each "next word"
  t = 1
  while (t < length(words)){
    all_mat[t, t+1] <- 1
    t = t + 1
  }
  
  #head(all_mat)

  
#Create a marix with all of the words listed as row names and only unique names listed as column names
  #get a list of unique words
  uniq_word = unique(words)
  #head(uniq_word)
  uni_col_mat = matrix(ncol = length(uniq_word), nrow = length(words),0)
  colnames(uni_col_mat) = uniq_word
  rownames(uni_col_mat) = words
  
  #head(uni_col_mat)
  
# For each unique name in columns' names find duplicated columns.Sum the values of each row for duplicate columns. 
# The result is a new matrix with the same number of rows with unique column names.
  for (i in 1:length(uniq_word)){
    #Use grep to get duplicated column names
    rep_col_In = grep(paste0("^",uniq_word[i],"$"),colnames(all_mat))
    # If there is only one column for a column name keep its values as is.
    if (length(rep_col_In) <= 1){
      col_name = colnames(all_mat)[rep_col_In]
      uni_col_mat[,col_name] = all_mat[,rep_col_In]
      #Else for each row sum the values for duplicated columns
    }else{
      col_name = colnames(all_mat)[rep_col_In][1]
      row_sum = rowSums(all_mat[,rep_col_In])
      uni_col_mat[,col_name] = row_sum
    }
  }
  
 # print(uni_col_mat)
  

#Create a marix of zeros with dimentions = number of unique words and unique words listed as row names and only unique names listed as column names
  uni_row_mat = matrix(ncol = length(uniq_word), nrow = length(uniq_word),0)
  colnames(uni_row_mat) = uniq_word
  rownames(uni_row_mat) = uniq_word
  
  # Repeat the above process for duplicated rows.
  #For each row with duplicated row names, add the values of each column.
  # The result is a matrix with unique rows and columns. Each number represent the frequency of column name to come after the row name.
  for (i in 1:length(uniq_word)){
    # Use grep to find duplicated rows
    rep_word = grep(paste0("^",uniq_word[i],"$"),rownames(uni_col_mat))
    # If there is only one row, keep the its column values as is.
    if (length(rep_word) <= 1){
      row = rownames(uni_col_mat)[rep_word]
      #print(row)
      uni_row_mat[row,]= uni_col_mat[rep_word,]
    # Else add the values of each column for duplicated rows
    }else{
      col_sum = colSums(uni_col_mat[rep_word,])
      name = rownames(uni_col_mat)[rep_word]
      #print(name)
      uni_row_mat[name[1],] = col_sum
    }
  }
#convert the frequencies into probablity fractions by deviding the non-zero values in each roow by  sum of the row.
  for (i in 1:dim(uni_row_mat)[1]) uni_row_mat[i, ] <- uni_row_mat[i, ] / sum(uni_row_mat[i, ])
  
  return(uni_row_mat)
}
```

This function generates new texts based on the probability matrix from the previous function
```{r}
predic = function(txt){
  #Comment this out if testing with another text
  txt = clean_txt(txt)
  words = unlist(strsplit(txt, " "))
  w = prob_mat(txt)
  
  dim_u = dim(w)[1]
  
  #select a random index in the range unique words
  current_index = sample(1:dim_u, 1)
  
  #Pick the first word based on the randomly selected index. 
  #We use the index to access the rowname
  current_word = as.character(rownames(w)[current_index])
  
  
  #Controler variable for the length of the output text
  book_len = 1
  
  while(book_len<length(words)){
    #Select the non_zero values from the row of first selected word
    # The column with non_zero values are words that could come after the current word
    non_zero = which(w[current_index,]!=0)
    
    #if there are more than one non_zero value, select one randomly, use prob the take the probablity values into account
    if(length(non_zero)>1){
      next_index = sample(non_zero,1, prob = w[current_index,which(w[current_index,]!=0)])
    } 
    #If there is only one non_zero value select that word as the next word
    else{
      next_index = non_zero
    }
    #Get the next word from column names
    next_word = as.character(colnames(w)[next_index])
    #Keep pasting the selected words after each other with a seperator
    current_word = paste(current_word,next_word, sep =' ')
    #Move to the next word
    current_index = next_index
    # The loop stops when the length of output text is equal to length of inpute text
    book_len=book_len+1
  }
  return(current_word)
}
```

Results
```{r}
test = predic(dt)
print(test)
```


